---
title: "CodeBook of Getting and Cleaning Data Course Project Outcome Data Set "
author: "SUN Lei"
date: "09/18/2014"
output: 
  html_document:
    keep_md: true
---

### Background Information

The data adopted is from "Human Activity Recognition Using Smartphones Dataset", which is the experiment outcome produced by

>Jorge L. Reyes-Ortiz, Davide Anguita, Alessandro Ghio, Luca Oneto in
>Smartlab - Non Linear Complex Systems Laboratory at
>DITEN - Università degli Studi di Genova.
>Via Opera Pia 11A, I-16145, Genoa, Italy.
>activityrecognition@smartlab.ws
>www.smartlab.ws

The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data. 

The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). Thereform there are 128 numbers record at each window. From each window, a vector of features was obtained by calculating variables from the time and frequency domain.

### Features/ Measurements
The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. These time domain signals (prefix 't' to denote time) were captured at a constant rate of 50 Hz. Then they were filtered using a median filter and a 3rd order low pass Butterworth filter with a corner frequency of 20 Hz to remove noise. Similarly, the acceleration signal was then separated into body and gravity acceleration signals (tBodyAcc-XYZ and tGravityAcc-XYZ) using another low pass Butterworth filter with a corner frequency of 0.3 Hz. 

Subsequently, the body linear acceleration and angular velocity were derived in time to obtain Jerk signals (tBodyAccJerk-XYZ and tBodyGyroJerk-XYZ). Also the magnitude of these three-dimensional signals were calculated using the Euclidean norm (tBodyAccMag, tGravityAccMag, tBodyAccJerkMag, tBodyGyroMag, tBodyGyroJerkMag). 

Finally a Fast Fourier Transform (FFT) was applied to some of these signals producing fBodyAcc-XYZ, fBodyAccJerk-XYZ, fBodyGyro-XYZ, fBodyAccJerkMag, fBodyGyroMag, fBodyGyroJerkMag. (Note the 'f' to indicate frequency domain signals). 

These signals were used to estimate variables of the feature vector for each pattern:  
'-XYZ' is used to denote 3-axial signals in the X, Y and Z directions.

- tBodyAcc-XYZ
- tGravityAcc-XYZ
- tBodyAccJerk-XYZ
- tBodyGyro-XYZ
- tBodyGyroJerk-XYZ
- tBodyAccMag
- tGravityAccMag
- tBodyAccJerkMag
- tBodyGyroMag
- tBodyGyroJerkMag
- fBodyAcc-XYZ
- fBodyAccJerk-XYZ
- fBodyGyro-XYZ
- fBodyAccMag
- fBodyAccJerkMag
- fBodyGyroMag
- fBodyGyroJerkMag

A set variables that were estimated from these signals but in this course we are only interested in ```mean(): Mean value``` and ```std(): Standard deviation```. Therefore the list of features are:
```{r echo = FALSE}
dfSetMean <- read.csv("step5_output2014-09-17 16:57:36.csv", nrows = 1, 
                      check.names = FALSE)
names(dfSetMean)[4:82]
```
Becasue the final outcome data is the average of each variable (feature/measurement) for each activity and each subject, a typical record in this outcome data set would be like:
```{r echo = FALSE}
dfSetMean
```
The above example shows that for Subject 20 in his/her layting activity, the average value of the X dimension body acceleration signals means ```tBodyAcc-mean()-X``` is 0.2682; his/her average value of the Z dimension body gyroscope signals standard deviations ```tBodyGyro-std()-Z``` is -0.4608; his/her average values of the magnitude of 3-dimension gravity acceleration signals ```tGravityAccMag-std()``` is -0.519; etc...



~ End ~

